### 2017-11-18 article/main.tex
Initial commit for article

\documentclass[conference]{IEEEtran}
\usepackage{blindtext, graphicx}

\usepackage[slovak]{babel}
\usepackage[utf8]{inputenc}
\usepackage{cite}

\usepackage{url}
\usepackage{amsmath}

\usepackage{algorithm}
\usepackage{algpseudocode}

\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Porovnanie času konvergencie siete medzi softvérovo riadenými sieťami a sieťami podporujúcimi protokol OSPF}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Peter Kaňuch}
\IEEEauthorblockA{Fakulta informatiky a informačných technológií\\
Slovenská Technická Univerzita\\
Slovenská republika, 841 04 Bratislava IV\\
Email: xkanuch@stuba.sk}\and
\IEEEauthorblockN{Nikolas Janec}
\IEEEauthorblockA{Fakulta informatiky a informačných technológií\\
Slovenská Technická Univerzita\\
Slovenská republika, 841 04 Bratislava IV\\
Email: xjanec@stuba.sk}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath

\end{abstract}

\begin{IEEEkeywords}

\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

\section{Úvod}

\section{}

\subsection{}

\subsection{}

%**********************************************************************************************************************
\section{}

\subsection{}

\subsection{}

\subsection{}

\section{}

\subsection{}

\subsection{}

\section{}

\section{}

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

\begin{thebibliography}{1}
\bibitem{}

\end{thebibliography}

\end{document}